---     
title: ""
author: ""
date: "July 16, 2019"
output:
  html_document: default
  pdf_document: default
---    
      
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
   
   
  

```{r message=FALSE, warning=FALSE}
list.of.packages <- c("expsmooth", "fpp2","fitdistrplus", "logspline","readxl", "xts","forecast", "fma","lmtest", "tseries","Quandl", "fpp","urca", "TSA","tidyverse","gdata","XLConnect")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages,dependencies=TRUE)
 
library('expsmooth')
library('fpp2')
library('fitdistrplus')
library('logspline')
library('xts')
library('forecast');
library('fma')
library('lmtest') 
library('tseries')
library('Quandl')
library('fpp');
library('urca')
library('TSA')
library('readxl')
library('tidyverse')
library('gdata')
#library('XLConnect')
#library('gdata')
#library('gdata')
```

### Combine the data from the 16 files into a single dataset and plot it:

```{r message=FALSE, warning=FALSE}
dataPath<-"~/Downloads/Time_Seriese/week4"
setwd(dataPath)
getwd()

file.list <- list.files(path = dataPath,pattern='*.xls')
df.list <- lapply(file.list, read_xls,sheet = 1, col_names = TRUE,skip = 1, range = cell_cols(c("C", "E")))

for (i in 1:length(df.list)){df.list[[i]]<-
cbind(df.list[[i]],datetime=as.Date(gsub("I-57-", "",gsub(".xls", "", file.list[i])),format="%Y-%B-%d"))
df.list[[i]] <- subset(df.list[[i]], select = -c(SI80) )
df.list[[i]] <- df.list[[i]][-c(1, 2, 27,28), ]
df.list[[i]]$datetime <- as.POSIXct(paste(df.list[[i]]$datetime,df.list[[i]]$Time),format="%Y-%m-%d %H:%M")
}
dataset<-do.call("rbind", df.list)
dataset<-dataset[, c("datetime", "I80E")]
dataset<-dataset[order(dataset$datetime),]
ts.dataset<-ts(dataset$I80E, frequency = 1)

autoplot(ts.dataset)

 
```
### Split the dataset into a  training dataset which includes 6/16/2013 -6/30/2013 samplesand a test dataset which includes 7/1/2013samples.Plot the ACF and PACF,and apply the Augmented Dickey-Fuller Test to check if the training dataset is stationary.
```{r message=FALSE, warning=FALSE}

myts.train.ds <- dataset[dataset$datetime<as.Date('2013-07-01'),2]
myts.train.ds<-as.numeric(myts.train.ds)
myts.test.ds  <- dataset[dataset$datetime>as.POSIXct('2013-06-30 23:00:00'),2]
myts.test.ds<-as.numeric(myts.test.ds)
# head(myts.train)
# tail(myts.train)
# head(myts.test)
myts.train<-ts(myts.train.ds, frequency = 24)
myts.test<-ts(myts.test.ds,start = c(7,1), frequency = 1)
fit <- tbats(myts.train)
seasonal <- !is.null(fit$seasonal)
print(paste('Seasonality = ',seasonal))
ggseasonplot(myts.train)

ggtsdisplay(myts.train)
adf.test(myts.train)


```

`Features:`

  `- seasonal pattern occurs.I used the tbats model.` 
  `   It will handle quarter seasonality and will automatically determine if a seasonal pattern is present, `
  `   I aslo used ggseasonplot and it show no Seasonality`
  
  `- series is stationary p value in the ADF test is 0.01`
  
  
### Build an Arima(p,d,q)model usingthe training dataset and R auto.arima()function.Change the values of p and q and determine the best model using AICc and BICvalues. AICc and BIC select the same model as the best model For each derived model
```{r message=FALSE, warning=FALSE}
(fit.arima <- auto.arima(myts.train, seasonal = FALSE))
(ESACF <- eacf(myts.train))
(EACF1 <- Arima(myts.train, order=c(2,0,1)))
(EACF2 <- Arima(myts.train, order=c(3,0,1)))
(EACF3 <- Arima(myts.train, order=c(7,0,1)))
(EACF4 <- Arima(myts.train, order=c(3,0,2)))
(EACF5 <- Arima(myts.train, order=c(2,0,3)))
(EACF6 <- Arima(myts.train, order=c(2,0,2)))
(Arima.AICs <- c(fit.arima=fit.arima$aic,EACF1=EACF1$aic,EACF2=EACF2$aic,EACF3=EACF3$aic,EACF4=EACF4$aic,EACF5=EACF5$aic,EACF6=EACF6$aic))
checkresiduals(EACF3)
```
`IEACF3 give better model than auto.arima`



### Build a day of the week seasonal Arima(p,d,q)(P,Q,D)s model using the training dataset and R auto.arima() function.


```{r message=FALSE, warning=FALSE}
myts.train.day <- ts(data = myts.train.ds,start = c(6,16), frequency = 168)
(fit.arima.day <- auto.arima(myts.train.day, seasonal = TRUE))
```


### Use the Arima(p,d,q)(P,Q,D)s model to forecast for July 1st(which is a Monday). Plot your result.
```{r message=FALSE, warning=FALSE}
forecast.day<-forecast(fit.arima.day, h= 24)
autoplot(forecast.day)

```

### Build a hour of the day seasonal Arima(p,d,q)(P,Q,D)s model using the training dataset and R auto.arima() function.

```{r message=FALSE, warning=FALSE}
myts.train.hour <- ts(data = myts.train.ds, frequency = 24)
(fit.arima.hour <- auto.arima(myts.train.hour, seasonal = TRUE))

```

### Use the Arima(p,d,q)(P,Q,D)s model to forecastfor July 1st (which is a Monday). Plot your result.

```{r message=FALSE, warning=FALSE}
forecast.hour<-forecast(fit.arima.hour, h= 24)
autoplot(forecast.hour)

```

### Compare the forecast of the models for July 1 8:00, 9:00, 17:00 and 18:00


```{r message=FALSE, warning=FALSE}

accuracy(forecast.day$mean[c(8,9,17,18)], myts.test[c(8,9,17,18)])
accuracy(forecast.hour$mean[c(8,9,17,18)], myts.test[c(8,9,17,18)])

autoplot(ts(myts.test[c(8,9,17,18)]))+ autolayer(ts(forecast.day$mean[c(8,9,17,18)]))+ggtitle("8,9,17,18 by day") 
autoplot(ts(myts.test[c(8,9,17,18)]))+ autolayer(ts(forecast.hour$mean[c(8,9,17,18)]))+ggtitle("8,9,17,18 by hour") 

autoplot(ts(myts.test))+ autolayer(ts(forecast.day$mean))+ggtitle("8,9,17,18 by day") 
autoplot(ts(myts.test))+ autolayer(ts(forecast.hour$mean))+ggtitle("8,9,17,18 by hour") 


```
`day of the week seasonal Arima is better than hour of the day seasonal Arima`
